---
title: "Probability of Genes"
author: "David Wong and Liam Carrigan"
date: "November 15, 2017"
output: html_document
---



```{r}
probcommon <- function(pool1, pool2, common, samp1size, samp2size, threshold)
{
  # set.seed(9999)
  # Numerically, Liam and I agreed that we required a number from Ricardo Frausto that was the quantity of genes in both pools that were already common between them.
  uncommon <- pool1 - common
  n1 <- 1:pool1
  n2 <- uncommon:(uncommon+pool2)
  
  commonl <- list()
  commoner <- 0
  
  for(j in 1:10000)
  {
    set.seed(j)
    a <- sample(n1, samp1size, replace = F)
    b <- sample(n2, samp2size, replace = F)
  
  for(i in 1:length(a))
    {
      for(k in 1:length(b))
      {
      if(a[i] == b[k])
        commoner = commoner + 1
      }
    }
  commonl[j] <- commoner
  commoner <- 0
  }
  results <- (unlist(commonl))
  pval <- length(which(commonl >= threshold))/10000
  expec <- mean(unlist(commonl))
  return(list(pval,expec, results))
}
```

Simulation 1
Probability of 40 genes being common between a sample of 108 selected from 8049 and 795 selected from 8432
with 6314 common genes
```{r}
sim1 <- probcommon(8049, 8432, 6314, 108, 795, 40)
```
1-10000 seeds:
P-value: 0.00
Mean:8.01



Simulation 2
Probability of 65 genes being common between a sample of 249 selected from 8384 and 920 selected from 8432 with 6459 common genes
```{r}
sim2 <- probcommon(8384, 8432, 6459, 249, 920, 65)
```
1-10000 seeds:
P-value: 0.00
Mean: 20.9


Simulation 3
Probility of 3 genes being common between a sample of 108 selected from 8049 and 920 selected from 8432 with 6314 common genes
```{r}
sim3 <- probcommon(8049, 8432, 6314, 108, 920, 3)
```
1-10000 seeds:
P-value:0.995
Mean: 9.27


Simulation 4
Probability of 3 genes being common between a sample of 249 selected from 8384 and 795 selected from 8432 with 6459 common genes
```{r}
sim4 <- probcommon(8384, 8432, 6459, 249, 795, 3)
```
1-10000 seeds:
P-value: 1.00
Mean: 18.1


```{r}
#Graphing
par(mfrow = c(2,2))
hist(sim1[[3]], main = "Sampling Distribution of Simulation 1", xlim = c(0, 42), breaks = 20, xlab = 'Number of Common Genes Out of 6314')
abline(v = 40, col = 'red', lwd = 2)

hist(sim2[[3]], main = "Sampling Distribution of Simulation 2", xlim = c(0, 72), breaks = 20, xlab = 'Number of Common Genes Out of 6459')
abline(v = 65, col = 'red', lwd = 2)

hist(sim3[[3]], main = "Sampling Distribution of Simulation 3", xlim = c(0, (max(sim3[[3]]) + 2)), breaks = 20, xlab = 'Number of Common Genes Out of 6314')
abline(v = 3, col = 'red', lwd = 2)

hist(sim4[[3]], main = "Sampling Distribution of Simulation 4", xlim = c(0, (max(sim4[[3]]) + 2)), breaks = 20, xlab = 'Number of Common Genes Out of 6459')
abline(v = 3, col = 'red', lwd = 2)
```



```{r, echo=FALSE}
#Creating table of means and p-values
means <- data.frame(matrix(c(sim1[[2]], sim2[[2]], sim3[[2]], sim4[[2]], sim1[[1]], sim2[[1]], sim3[[1]], sim4[[1]]), nrow = 2, byrow = TRUE))
colnames(means) <- c("Simulation 1", "Simulation 2", "Simulation 3", "Simulation 4")
rownames(means) <- c("Means", "P-Values")


```

--------------------------------------------------------------------------



  